---
title: "Untitled"
author: "Niko Partanen"
date: "11/17/2017"
output: md_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


# metadata-concepts

This repository contains a listing of metadata values used in [IKDP](https://github.com/langdoc/IKDP) and [IKDP-2](https://github.com/langdoc/IKDP-2) language documentation projects, and the values are mapped between CMDI, project's internal Filemaker Pro database, and the terms Niko Partanen attempts to use in project's more or less internal R scripts and functions.

The discussion about standards seem to circulate often around **the format structure**, which in many ways is the least important, most mechanical, part of the question. It is more important to think what kind of data and values we have and need. It seems that these values eventually are rather similar across projects, although the exact implementation and structures would differ. Even more essentially, it seems there is kind of a **core** for possible information, without which the data is almost unusable to start with. It is essential to know **where** something was done, **when** and **with who**, but then there are details that are arguably less crucial. One way to think about this is that there are some values we just have to have and can have for every recording we work with, so there is some kind of a basic minimum on which we can compare every item, but then there are attributes which exclude some sessions as we don't have sufficient metadata.

So more recording sessions we have to exclude from comparison because the metadata field in question is lacking, more peripheral that field has to be. Someone can argue it is super-important, but if nobody has ever collected it, then it probably is not. Another example is that for M.A. Castrén's data (1844) we can't really specify the participants or locations, whereas for bit newer data such as that of Fokos-Fuchs's (1913) we generally can do that, up to a certain degree, so this sets kind of a non-arbitrary boundary between these two datasets.

It is obvious that in different environments there are needs to call things different. What would be in a presentation or on a website `English title` is in database something like `title-en` and in R dataframe `title_en`. All these refer to the same concept, and the idea here is to map all these concepts and their variants to one another. Similarly different concepts have to be mapped to their translations, and ideally everything would be done in a manner where each piece of information would be stored only once.

```{r, echo=FALSE}
suppressPackageStartupMessages(library(tidyverse))
read_csv('concept-mapping.csv', col_types = 'ccc', na = " ", quoted_na = TRUE) %>%
  knitr::kable(format = "markdown")
```

## Tasks

- Setting up Travis CI to compile README.Rmd automatically
- Add new values and mappings
- There has to be some way to indicate that things like **session date** is actually stored only in the session name.
    - For example, there are conventions such as session name has fields: `{language}_{variety}{year}{month}{date}{nth-recording of a day}`
    - So session date value actually maps into `{year}{month}{date}` complex in the session name… Maybe not ideal, but if you have that value somewhere in the database, then you have to maintain two distinct values, and having it in filename just makes browsing files million times easier!

## Notes about fields to be added

Table below created with Dmitriy Levchenko. It mainly contains fields we need to get into our system in Syktyvkar.

- participant
    - participant-id
    - participant-name
        - type: first, middle, patronym, last
        - language: kpv, rus
    - participant-place
        - type: birth, current, stayed
        - period: start time, end time
    - participant-photo
    - participant-gender
- session
    - Id
        - location
    - time
        - period: start time, end time
    - tematika
    - title
        - language: kpv, rus, eng
    - status
    - filename
    - comment
        - language: kpv, rus, eng
    - project
    - source
    - description
        - language: kpv, rus, eng
    - language
- connection
    - Id informant
    - Id zaps
    - Rol
- location
    - id
    - name
        - language: kpv, rus
    - lat
    - lon

As a relational database, same structure, approximately, can be expressed like this:

- Participants
    - Id
    - Name
    - Family
    - Patronym
    - Year
    - Birthplace id
    - Place of residence id
    - Picture (id?)
    - Gender
- Session
    - Id
    - Recording place id
    - Time
    - Topics
    - Title-ru
    - Title-en
    - Status (segmented, transcribed etc.)
    - Session name
    - Comments
    - Project
    - Source (published somewhere etc.)
    - Description
- Connections
    - Id participant
    - Id session
    - Role
- Places
    - Id
    - Name rus
    - Name kpv
    - Coordinates

